Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2017-09-06 Ben_Manuscripts/Paper1/Outline.tex
response to feedback on results and discussion

  The initial distance between layers can influence the approach towards
  equilibrium. 
  \begin{itemize}
  	\item Equilibrated systems built according to the 3.7 \angstrom 
	layer spacing implied by R-pi are characterized by a defined, cylindrical and
	open pore structure.
	\item Aromatic rings are arranged in a helical conformation.%BJC: could potentially support with a figure. Mentioning the helical arrangmentment might makes things confusing though
        %MRS: 'basin' is not that common a term, and also only makes sense if one explicitly states what 
        %MRS: the basin is with respect to. 
        We will refer to this large set of configurations, with an open helical pore, as Basin A (Figure~\ref{fig:abasin}).
        %\item We will refer to this system as
        %MRS: maybe Basin O and Basin C for open and closed?
	\item Simulations of systems built with layers stacked greater than 4 
	\angstrom apart results in a pore structure characterized by high radial 
	disorder, while still maintaining partitioning between hydrophobic and 
	hydrophilic regions.
        %MRS: see above for clarifying ``basins''.
	\item This will be called Basin B (Figure~\ref{fig:bbasin}).
        %MRS: I don't think enough information has been presented to the reader yet for this to be 'apparent'.
	\item It is apparent that this LLC membrane may exist in at least two metastable states.
	\item The distinct difference in pore structure exhibited by each phase will 
	likely lead to different transport mechanisms.
	\item Understanding which phase exists experimentally is necessary in order
	to appropriately study the system.  %MRS: remind people why.
  \end{itemize}

@@

  We varied the relative interlayer orientation between sandwiched and 
  parallel-displaced based on our knowledge of the stability of the two
  pi-stacking modes.
  \begin{itemize}
  	\item We generated simulated X-ray diffraction patterns from the highly
	restrained initial configurations created during equilibration
	\item The patterns establish a difference between the two stacking 
	modes (Figure~\ref{fig:XRDrestrained}).
	\item In each pattern, R-alkanes is present at a distance of $\approx
	1.5$ \angstrom \textsuperscript{-1} (4.2 \angstrom in real space).
  	\item R-pi is also present in each pattern, although it appears to 
	intersect R-alkanes because the spacing between aromatic rings is 
	similar to the alkane chain packing.
        %MRS: for above, may need to clarify that this is in simulation, which is different than experiment.
	\item The difference in aromatic ring spacing between experiment and
	simulation is likely a result of the inability of GAFF to properly handle
	aromatic interactions. %MRS: provide citations, some mechanistic reasoning.
	\item The sandwiched configuration shows R-spots in the expected location.
	\item A faint reflection is present in the location of R-helix %BJC: but it fades upon further simulation.
	\item The parallel displaced configuration contains two lines of high 
	intensity with maximum intensity occurring where they intersect the alkane
	chain region due to constructive interference between the two features.
	\item The lines are located where one would expect to see R-helix and, 
	perhaps coincidentally, the intersection with R-alkanes occurs where one 
	would expect to see R-spots.
	%BJC: might save the following sentence for equilibrium trajectories
	\item As the simulation is progressed, the full line begins to fade 
	(Figure~\ref{fig:XRDsim}). Most significantly, the line is no longer present
	where R-helix exists.
  \end{itemize}

  \begin{figure}[!ht]
	\centering
	\begin{subfigure}{0.475\textwidth}
		\hspace{-1.2cm}
		\centering
		\includegraphics[width=\textwidth,valign=t]{rzlayeredrestrained.png}
		\caption{}\label{fig:rzplayeredrestrained}
	\end{subfigure}
	\begin{subfigure}{0.475\textwidth}
		\hspace{-1.2cm}
		\centering
		\includegraphics[width=\textwidth,valign=t]{rzoffsetrestrained.png}
		\caption{}\label{fig:rzoffsetrestrained}
	\end{subfigure}
	\caption{(a) Simulated X-ray diffraction of a sandwiched configuration
	with restraints placed on aromatic carbons shows all major features
	present in experimental WAXS. Near solid lines at constant z are a result of 
	the highly ordered aromatic rings. (b) Simulated X-ray diffraction of a similarly
	restrained parallel displaced configuration may also contain all
	major experimental WAXS features. One can argue that R-spots is not present
	but it is difficult to distinguish because of its intersection with the solid
	R-helix line. 
        %MRS: Longer simulations are necessary to determine which structure
	is the best match to experiment
        %MRS: we probably want to avoid saying this (above) if we can.
}\label{fig:XRDrestrained}
  \end{figure}

  %MRS: one thought: maybe organize it a different way -- it's not the disordered (and go through the anlysis there), and then separately show that it's more likely the sandwich.  Something to think about.
  %MRS: anyway, this is the key part of the paper, so clarity and organization important.
  %MRS: make sure the figures are comparable.  Could someone look at the figures, and see immediately what you are explaining -- for example, experiment and fully equilibrated disordered, sandwich and parallel displaced all in a row?
  Full comparison of experimental 2D WAXS with simulated X-ray diffraction
  patterns produced from equilibrated MD trajectories shows the most consistency
  with the sandwiched configuration in Basin A.
  \begin{itemize}
  	\item Systems were built in both the parallel displaced and sandwiched 
	configurations with an intial layer spacing of 3.7 \angstrom.
	\item A third system was created by stacking layers in the sandwiched 
	configuration 5 \angstrom apart in order to guide it towards Basin B.
	% BJC: I've done an additional offset disordered system which goes to Basin B but I think that will confuse the point
	\item For ease of reference, the third system will be referred to as the disordered system. 
%BJC: omit and call it basin B
%MRS or just be consistent from the beginning.  Or call Basins O and D          
	\item The three systems were equilibrated according to our procedure with
	NPT simulations of greater than 400 ns.
	\item Simulated diffration patterns were generated using portions of the
	trajectory after equilibration.
%MRS: something like ``we assume the simulation has equilibrated when the distance . . . ``        
	\item Equilibration was detected when the distance between pores and the 
	membrane thickness stopped changing.
	\item Simulated diffraction patterns for all three structures are shown in 
	Figure ~\ref{fig:XRDsim}. 
	%BJC: maybe break into a separate paragraph starting here
	\item The equilibrated parallel displaced structure exhibits R-alkanes and
	R-pores in the expected locations. 
	\item Due to low resolution, making out the individual spots of R-pores is
	challenging and can be validated upon full spherical integration of the 3D 
	structure factor. However, the same information can be extracted by measuring 
	the pore spacing as described earlier.
	\item R-pi is also present, intersecting R-alkanes at a lower q value than
	in experiment. The rings prefer to stack $\approx 4.1 \angstrom$ apart as 
	opposed to 3.7 \angstrom.
	\item R-helix and R-spots are also both missing.
	\item The line which intersected both of these features in the restrained 
	simulations faded upon equilibration which suggests that monomers in the 
	parallel-displaced configuration may not give rise to the experimental WAXS 
	pattern seen.
	\item Simulated XRD of the sandwiched configuration contains all features 
	except for R-helix.
	\item Most notably, R-spots appears in the expected location, which suggests 
	that there is something intrinsic to the sandwiched packing that gives rise 
	to such features. 
	\item Finally, the disordered structure exhibits R-alkanes and R-pores but 
	R-helix, R-pi and R-spots are not present.
  \end{itemize}

@@

  \begin{wrapfigure}{R}{0.4\textwidth}
        \centering
        \includegraphics[width=0.4\textwidth]{tilt.png}
        \caption{The average angle between alkane chains and the xy plane is nearly zero degrees}\label{fig:tilt}
        %MRS: probably could just report this in text, put figure in supporting.
  \end{wrapfigure}

  The spots that appear in the simulated XRD pattern of the sandwiched
  conformation are a result of the way alkane tails pack together.
  \begin{itemize}
  	\item Previously, the spots in the diffraction pattern had been explained 
	as the product of tilted alkane chains.
	\item We measured the tilt angle of the alkane chains and showed that our 
	system equilibrates to an average tilt angle close to zero degrees (Figure~\ref{fig:tilt}). 
	\item To understand the origin of the spots, we determined which atoms gave rise to the feature.
	\item By removing atoms from the trajectory and simulating a diffraction 
	pattern, we were able to isolate the cause of the spots to the tails (Figure~\ref{fig:tails_rzplot}).
	\item Since the tails are mostly flat, we plotted the centroids of the 
	tails and measured the angle between each centroid and its nearest neighbors
	with respect to the plane of the membrane (Figure~\ref{fig:centroids}).
	\item The distribution of these angles is consistent with the location of 
	the spots (Figure~\ref{fig:angle_distribution}).
	\item We reason that monomer tails stacked closely in the sandwiched conformation
	are forced to splay apart and pack in between in each other which creates a
	nearly hexagonal array of packed tails. 
        %MRS: hmm.  and this doens't happen in parallel displaced? I thought it did.
        %thought it did before. Would be good to explore the equivalent of figure 9 to this, because that comparison would make it clearer what the structural reason is for the R-spots.
  \end{itemize}

@@

                \caption{}\label{fig:thickness_step}
        \end{subfigure}
        \begin{subfigure}[b]{0.325\textwidth}
                \centering
                \includegraphics[width=\textwidth]{order_layered.png}
                \caption{}\label{fig:order_step}
        \end{subfigure}
        %MRS: Are (a) and (b) from the nonequilibrium short run?
        %That's not clear. Not clear it's possible to state that it's
        %the ``right'' disordered basin if from the end of a short heating run.

@@

  %MRS: which model? Be specific as to which of the above structural ensembles was used.
  The model gives reasonable estimates of ionic conductivity when compared to 
  experiment.
  \begin{itemize}
  	\item Calculated values of ionic conductivity obtained using the Nernst 
	Einstein relation and Collective Diffusion model are compared in 
	Table~\ref{table:conductivity}.
	\item The two methods agree with each other within error, although the 
	uncertainty obtained using the Collective Diffusion model is much higher.
	\item Much longer simulations are needed to lower the uncertainty. % TODO: Run Basin B out longer at same conditions as A
	\item Collective diffusion calculations were generated from 
	500 ns simulations.
	\item Our calculated values would benefit from longer simulations. 
	\item For this reason we will likely only use the Nernst Einstein relation
	in future calculations of this type. 
	\item Interestingly, Basin B has a higher ionic conductivity than Basin A.
	\item We hypothesize that conductivity is enhanced in Basin B due to a 
	higher sodium ion diffusivity.
	\item Transport of sodium is likely facilitated by the homogeneity of Basin B.
	Sodium ions have less sites to move to in Basin A.
        % total number of sites is the same in both basins, correct? -- do you mean that in basin B, there are more nearby sites?
	\item There is currently no experimental evidence of this trend. Maybe Xund will find something
	\item In both cases, our calculated values for Basin A are higher than the 
	experimental values, as expected. 
	\item Some of the discrepancy is likely a result of using an imperfect forcefield. 
	\item However, the real system, although mostly aligned and straight, has a 
	distribution of azimuthal angles, meaning that the pores have a degree of 
	tortuosity which lowers the effective ionic conductivity of the bulk membrane. 
	\item The ordering from isotropic to mostly aligned mesophases showed an 85 
	fold increase in ionic conductivity. We would expect additional gains in a 
	perfectly aligned system.
  \end{itemize}

@@

>>>>

  The initial distance between layers can influence the approach towards
  equilibrium. 
  \begin{itemize}
  	\item Our choice of distance between layers results in long-term stability of 
	different configurations indicating that we have found more than one metastable
	free energy basin.
  	\item Equilibrated systems built according to the 3.7 \angstrom 
	layer spacing implied by R-pi are characterized by a defined, cylindrical and
	open pore structure.
	\item Aromatic rings are arranged in a helical conformation.%BJC: could potentially support with a figure. Mentioning the helical arrangmentment might makes things confusing though
        %MRS: 'basin' is not that common a term, and also only makes sense if one explicitly states what 
        %MRS: the basin is with respect to. 
	%BJC: added sentence to beginning of paragraph
        \item We will refer to this large set of configurations, with an open helical pore, as Basin A (Figure~\ref{fig:abasin}).
        %\item We will refer to this system as
        %MRS: maybe Basin O and Basin C for open and closed?
	%BJC: That works but 'capital oh' might get confused with zero
	\item Simulations of systems built with layers stacked greater than 4 
	\angstrom apart results in a pore structure characterized by high radial 
	disorder, while still maintaining partitioning between hydrophobic and 
	hydrophilic regions.
        %MRS: see above for clarifying ``basins''.
	\item This will be called Basin B (Figure~\ref{fig:bbasin}).
        %MRS: I don't think enough information has been presented to the reader yet for this to be 'apparent'.
	\item This LLC membrane may exist in at least two metastable states.
	\item The distinct difference in pore structure exhibited by each phase will 
	likely lead to different transport mechanisms.
	\item Because the pore structure will vary between each Basin, understanding 
	which phase exists experimentally is necessary in order to ensure we are 
	studying the system which actually dominates.  %MRS: remind people why.
  \end{itemize}

@@

  We varied the initial relative interlayer orientation between sandwiched and 
  parallel-displaced based on our knowledge of the stability of the two
  pi-stacking modes.
  \begin{itemize}
  	\item We generated simulated X-ray diffraction patterns using simulation trajectories from the highly
	restrained initial configurations created during equilibration
	\item The simulated patterns establish a difference between the two stacking 
	modes (Figure~\ref{fig:XRDrestrained}).
	\item In each pattern, R-alkanes is present at a distance of $\approx
	1.5$ \angstrom \textsuperscript{-1} (4.2 \angstrom in real space).
  	\item R-pi is also present in each pattern, although it appears to 
	intersect R-alkanes because the spacing between aromatic rings is 
	similar to the alkane chain packing.
        %MRS: for above, may need to clarify that this is in simulation, which is different than experiment.
	%BJC: okay added some more specific language
	\item The difference in aromatic ring spacing between experiment and
	simulation is likely a result of the inability of GAFF to properly handle
	aromatic interactions. %MRS: provide citations, some mechanistic reasoning.
	\item The sandwiched configuration shows R-spots in the expected location.
	\item A faint reflection is present in the location of R-helix %BJC: but it fades upon further simulation.
	\item The parallel displaced configuration contains two lines of high 
	intensity with maximum intensity occurring where they intersect the alkane
	chain region due to constructive interference between the two features.
	\item The lines are located where one would expect to see R-helix and, 
	perhaps coincidentally, the intersection with R-alkanes occurs where one 
	would expect to see R-spots.
	%BJC: might save the following sentence for equilibrium trajectories
	\item As the simulation is progressed, the full line begins to fade 
	(Figure~\ref{fig:XRDsim}). Most significantly, the line is no longer present
	where R-helix exists.
  \end{itemize}

  \begin{figure}[!ht]
	\centering
	\begin{subfigure}{0.475\textwidth}
		\hspace{-1.2cm}
		\centering
		\includegraphics[width=\textwidth,valign=t]{rzlayeredrestrained.png}
		\caption{}\label{fig:rzplayeredrestrained}
	\end{subfigure}
	\begin{subfigure}{0.475\textwidth}
		\hspace{-1.2cm}
		\centering
		\includegraphics[width=\textwidth,valign=t]{rzoffsetrestrained.png}
		\caption{}\label{fig:rzoffsetrestrained}
	\end{subfigure}
	\caption{(a) Simulated X-ray diffraction of a sandwiched configuration
	with restraints placed on aromatic carbons shows all major features
	present in experimental WAXS. Near solid lines at constant z are a result of 
	the highly ordered aromatic rings. (b) Simulated X-ray diffraction of a similarly
	restrained parallel displaced configuration may also contain all
	major experimental WAXS features. One can argue that R-spots is not present
	but it is difficult to distinguish because of its intersection with the solid
	R-helix line. 
        %MRS: Longer simulations are necessary to determine which structure
	% is the best match to experiment
        %MRS: we probably want to avoid saying this (above) if we can.
}\label{fig:XRDrestrained}
  \end{figure}

  %MRS: one thought: maybe organize it a different way -- it's not the disordered (and go through the anlysis there), and then separately show that it's more likely the sandwich.  Something to think about.
  %MRS: anyway, this is the key part of the paper, so clarity and organization important.
  %MRS: make sure the figures are comparable.  Could someone look at the figures, and see immediately what you are explaining -- for example, experiment and fully equilibrated disordered, sandwich and parallel displaced all in a row?
  Full comparison of experimental 2D WAXS with simulated X-ray diffraction
  patterns produced from equilibrated MD trajectories shows the most consistency
  with the sandwiched configuration in Basin A.
  \begin{itemize}
  	\item Systems were built in both the parallel displaced and sandwiched 
	configurations with an intial layer spacing of 3.7 \angstrom.
	\item A third system was created by stacking layers in the sandwiched 
	configuration 5 \angstrom apart in order to guide it towards Basin B.
	% BJC: I've done an additional offset disordered system which goes to Basin B but I think that will confuse the point
	\item For ease of reference, the third system will be referred to as the disordered system. 
%BJC: omit and call it basin B
%MRS or just be consistent from the beginning.  Or call Basins O and D          
	\item The three systems were equilibrated according to our procedure with
	NPT simulations of greater than 400 ns.
	\item Simulated diffration patterns were generated using portions of the
	trajectory after equilibration.  
	\item We assume the simulation has equilibrated when the distance between 
	pores and the membrane thickness stopped changing.
	\item Simulated diffraction patterns for all three structures are shown in 
	Figure ~\ref{fig:XRDsim}. 
	%BJC: maybe break into a separate paragraph starting here
	\item The equilibrated parallel displaced structure exhibits R-alkanes and
	R-pores in the expected locations. 
	\item Due to low resolution, making out the individual spots of R-pores is
	challenging and can be validated upon full spherical integration of the 3D 
	structure factor. However, the same information can be extracted by measuring 
	the pore spacing as described earlier.
	\item R-pi is also present, intersecting R-alkanes at a lower q value than
	in experiment. The rings prefer to stack $\approx 4.1 \angstrom$ apart as 
	opposed to 3.7 \angstrom.
	\item R-helix and R-spots are also both missing.
	\item The line which intersected both of these features in the restrained 
	simulations faded upon equilibration which suggests that monomers in the 
	parallel-displaced configuration may not give rise to the experimental WAXS 
	pattern seen.
	\item Simulated XRD of the sandwiched configuration contains all features 
	except for R-helix.
	\item Most notably, R-spots appears in the expected location, which suggests 
	that there is something intrinsic to the sandwiched packing that gives rise 
	to such features. 
	\item Finally, the disordered structure exhibits R-alkanes and R-pores but 
	R-helix, R-pi and R-spots are not present.
  \end{itemize}

@@

% Supplemental VVV
  \begin{wrapfigure}{R}{0.4\textwidth}
        \centering
        \includegraphics[width=0.4\textwidth]{tilt.png}
        \caption{The average angle between alkane chains and the xy plane is nearly zero degrees}\label{fig:tilt}
        %MRS: probably could just report this in text, put figure in supporting.
  \end{wrapfigure}

  The spots that appear in the simulated XRD pattern of the sandwiched
  conformation are a result of the way alkane tails pack together.
  \begin{itemize}
  	\item Previously, the spots in the diffraction pattern had been explained 
	as the product of tilted alkane chains.
	\item We measured the tilt angle of the alkane chains and showed that our 
	system equilibrates to an average tilt angle close to zero degrees (Figure~\ref{fig:tilt}). 
	\item To understand the origin of the spots, we determined which atoms gave rise to the feature.
	\item By removing atoms from the trajectory and simulating a diffraction 
	pattern, we were able to isolate the cause of the spots to the tails (Figure~\ref{fig:tails_rzplot}).
	\item Since the tails are mostly flat, we plotted the centroids of the 
	tails and measured the angle between each centroid and its nearest neighbors
	with respect to the plane of the membrane (Figure~\ref{fig:centroids}).
	\item The distribution of these angles is consistent with the location of 
	the spots (Figure~\ref{fig:angle_distribution}).
	\item We reason that monomer tails stacked closely in the sandwiched conformation
	are forced to splay apart and pack in between in each other which creates a
	nearly hexagonal array of packed tails. 
        %MRS: hmm.  and this doens't happen in parallel displaced? I thought it did.
        %thought it did before. Would be good to explore the equivalent of figure 9 to this, because that comparison would make it clearer what the structural reason is for the R-spots.
	%BJC: There aren't spots in parallel displaced and the distribution does not have the spikes that the sandwiched configuration does. At this point in the paper, I've tried to say that parallel displaced configuration is not consistent with experiment and I focus the rest of the paper on the sandwiched configuration
  \end{itemize}

@@

                \caption{}\label{fig:thickness_step}
        \end{subfigure}
        \begin{subfigure}[b]{0.325\textwidth}
                \centering
                \includegraphics[width=\textwidth]{order_layered.png}
                \caption{}\label{fig:order_step}
        \end{subfigure}
        %MRS: Are (a) and (b) from the nonequilibrium short run?
        %That's not clear. Not clear it's possible to state that it's
        %the ``right'' disordered basin if from the end of a short heating run.
	%BJC: They are from the short run, but as I comment above, I need to 
	% replace them with the correct figures

@@

  %MRS: which model? Be specific as to which of the above structural ensembles was used.
  The model gives reasonable estimates of ionic conductivity when compared to 
  experiment.
  \begin{itemize}
  	\item Calculated values of ionic conductivity obtained using the Nernst 
	Einstein relation and Collective Diffusion model are compared in 
	Table~\ref{table:conductivity}.
	\item The two methods agree with each other within error, although the 
	uncertainty obtained using the Collective Diffusion model is much higher.
	\item Much longer simulations are needed to lower the uncertainty. % TODO: Run Basin B out longer at same conditions as A
	\item Collective diffusion calculations were generated from 
	500 ns simulations.
	\item Our calculated values would benefit from longer simulations. 
	\item For this reason we will likely only use the Nernst Einstein relation
	in future calculations of this type. 
	\item Interestingly, Basin B has a higher ionic conductivity than Basin A.
	\item We hypothesize that conductivity is enhanced in Basin B due to a 
	higher sodium ion diffusivity.
	\item Transport of sodium is likely facilitated by the homogeneity of Basin B.
	Sodium ions have less nearby sites to move to in Basin A.
	\item There is currently no experimental evidence of this trend. Maybe Xunda will find something
	\item In both cases, our calculated values for Basin A are higher than the 
	experimental values, as expected. 
	\item Some of the discrepancy is likely a result of using an imperfect forcefield. 
	\item However, the real system, although mostly aligned and straight, has a 
	distribution of azimuthal angles, meaning that the pores have a degree of 
	tortuosity which lowers the effective ionic conductivity of the bulk membrane. 
	\item The ordering from isotropic to mostly aligned mesophases showed an 85 
	fold increase in ionic conductivity. We would expect additional gains in a 
	perfectly aligned system.
  \end{itemize}

@@

  \section*{Conclusion}
